\pkg{cv}: An \proglang{R} Package for Cross-Validation of Regression Models

### 2024-01-03 scratch/notes_john/JSS-article/JSS-article-reduced.Rmd
added JSS-article-reduced.Rmd (preliminary attempt)

## Logistic regression for the Mroz data

The \code{Mroz} data set from the \pkg{carData} package [associated with @FoxWeisberg:2019] has been used by several authors to illustrate binary logistic regression; see, in particular @FoxWeisberg:2019. The data were originally drawn from the U.S. Panel Study of Income Dynamics and pertain to married women. Here are a few cases in the data set:
```{r Mroz-data}
data("Mroz", package="carData")
head(Mroz, 3)
tail(Mroz, 3)
```
The response variable in the logistic regression is \code{lfp}, labor-force participation, a factor coded \code{"yes"} or \code{"no"}. The remaining variables are predictors:

* \code{k5}, number of children 5 years old of younger in the woman's household;
* \code{k618}, number of children between 6 and 18 years old;
* \code{age}, in years;
* \code{wc}, wife's college attendance, \code{"yes"} or \code{"no"};
* \code{hc}, husband's college attendance;
* \code{lwg}, the woman's log wage rate if she is employed, or her *imputed* wage rate, if she is not [a variable that @FoxWeisberg:2019 show is problematically defined]; and
* \code{inc}, family income, in $1000s, exclusive of wife's income.

We use the \code{glm()} function to fit a binary logistic regression to the \code{Mroz} data:
```{r Mroz-logistic-regresion}
m.mroz <- glm(lfp ~ ., data=Mroz, family=binomial)
summary(m.mroz)

BayesRule(ifelse(Mroz$lfp == "yes", 1, 0), 
          fitted(m.mroz, type="response"))
```
In addition to the usually summary output for a GLM, we show the result of applying the \code{BayesRule()} function from the \pkg{cv} package to predictions derived from the fitted model. Bayes rule, which predicts a "success" in a binary regression model when the fitted probability of success [i.e., $\phi = \Pr(y = 1)$] is $\widehat{\phi} \ge .5$ and a "failure" if $\widehat{\phi} < .5$.[^BayesRule] The first argument to \code{BayesRule()} is the binary {0, 1} response, and the second argument is the predicted probability of success. \code{BayesRule()} returns the proportion of predictions that are *in error*, as appropriate for a "cost" function.

[^BayesRule]: \code{BayesRule()} does some error checking; \code{BayesRule2()} is similar, but omits the error checking, and so can be faster for large problems.

The value returned by \code{BayesRule()} is associated with an "attribute" named \code{"casewise loss"} and set to \code{"y != round(yhat)"}, signifying that the Bayes rule CV criterion is computed as the mean of casewise values, here 0 if the prediction for a case matches the observed value and 1 if it does not (signifying a prediction error). The \code{mse()} function for numeric responses is also calculated as a casewise average. Some other criteria, such as the median absolute error, computed by the \code{medAbsErr()} function in the \pkg{cv} package, aren't averages of casewise components. The distinction is important because, to our knowledge, the statistical theory of cross-validation, for example, in @DavisonHinkley:1997, @BatesHastieTibshirani:2023, and @ArlotCelisse:2010, is developed for CV criteria like MSE that are means of casewise components. As a consequence, we limit computation of bias adjustment and confidence intervals (see below) to criteria that are casewise averages.

In this example, the fitted logistic regression incorrectly predicts 31% of the responses; we expect this estimate to be optimistic given that the model is used to "predict" the data to which it is fit.

The \code{"glm"} method for \code{cv()} is largely similar to the \code{"lm"} method, although the default algorithm, selected explicitly by \code{method="exact"}, refits the model with each fold removed (and is thus equivalent to \code{method="naive"} for \code{"lm"} models). For generalized linear models, \code{method="Woodbury"} or (for LOO CV) \code{method="hatvalues"} provide approximate results (see the last section of the vignette for details):

```{r cv-Mroz-10-fold}
cv(m.mroz, criterion=BayesRule, seed=248)

cv(m.mroz, criterion=BayesRule, seed=248, method="Woodbury")
```
To ensure that the two methods use the same 10 folds, we specify the seed for \proglang{R}'s random-number generator explicitly; here, and as is common in our experience, the \code{"exact"} and \code{"Woodbury"} algorithms produce nearly identical results. The CV estimates of prediction error are slightly higher than the estimate based on all of the cases.

The printed output includes a 95% confidence interval for the bias-adjusted Bayes rule CV criterion. @BatesHastieTibshirani:2023 show that these confidence intervals are unreliable for models fit to small samples, and by default \code{cv()} computes them only when the sample size is 400 or larger and when the CV criterion employed is an average of casewise components, as is the case for Bayes rule. See the final section of the vignette for details of the computation of confidence intervals for bias-adjusted CV criteria.

Here are results of applying LOO CV to the Mroz model, using both the exact and the approximate methods:
```{r cv-Mroz-LOO}
cv(m.mroz, k="loo", criterion=BayesRule)

cv(m.mroz, k="loo", criterion=BayesRule, method="Woodbury")

cv(m.mroz, k="loo", criterion=BayesRule, method="hatvalues")
```
To the number of decimal digits shown, the three methods produce identical results for this example.

As for linear models, we report some timings for the various \code{cv()` methods of computation in LOO CV as well as for the \code{cv.glm()} function from the \pkg{boot} package (which, recall, refits the model with each case removed, and thus is comparable to \code{cv()} with \code{method="exact"}):
```{r glm.timings, cache=TRUE}
microbenchmark::microbenchmark(
  hatvalues=cv(m.mroz, k="loo", criterion=BayesRule, method="hatvalues"),
  Woodbury=cv(m.mroz, k="loo", criterion=BayesRule, method="Woodbury"),
  exact=cv(m.mroz, k="loo", criterion=BayesRule),
  cv.glm=boot::cv.glm(Mroz, m.mroz,
               cost=BayesRule),
  times=10)
```
There is a substantial time penalty associated with exact computations.

@@

## Example: Crossed random effects

Crossed random effects arise when the structure of the data aren't strictly hierarchical. Nevertheless, crossed and nested random effects can be handled in much the same manner, by refitting the mixed-effects model to the data with a fold of clusters or cases removed and using the refitted model to predict the response in the removed fold.

We'll illustrate with data on pig growth, introduced by @DiggleLiangZeger:1994 [Table 3.1]. The data are in the \code{Pigs} data frame in the \pkg{cv} package:
```{r pigs}
head(Pigs, 9)
head(xtabs(~ id + week, data=Pigs), 3)
tail(xtabs(~ id + week, data=Pigs), 3)
```
Each of 48 pigs is observed weekly over a period of 9 weeks, with the weight of the pig recorded in kg. The data are in "long" format, as is appropriate for use with the \code{lmer()} function in the \pkg{lme4} package. The data are very regular, with no missing cases.

The following graph, showing the growth trajectories of the pigs, is similar to Figure 3.1 in @DiggleLiangZeger:1994; we add an overall least-squares line and a loess smooth, which are nearly indistinguishable:
```{r pigs-graph}
#| out.width = "60%",
#| fig.height = 6,
#| fig.cap = "Growth trajectories for 48 pigs, with overall least-squares line (sold blue) and loess line (broken magenta)."
plot(weight ~ week, data=Pigs, type="n")
for (i in unique(Pigs$id)){
  with(Pigs, lines(x=1:9, y=Pigs[id == i, "weight"],
                   col="gray"))
}
abline(lm(weight ~ week, data=Pigs), col="blue", lwd=2)
lines(with(Pigs, loess.smooth(week, weight, span=0.5)),
      col="magenta", lty=2, lwd=2)
```
The individual "growth curves" and the overall trend are generally linear, with some tendency for variability of pig weight to increase over weeks (a feature of the data that we ignore in the mixed model that we fit to the data below).

The \proglang{Stata} mixed-effects models manual proposes a model with crossed random effects for the \code{Pigs} data [@Stata:2023 page 37]:

> [S]uppose that we wish to fit
$$
\mathrm{weight}_{ij} = \beta_0 + \beta_1 \mathrm{week}_{ij} + u_i + v_j + \varepsilon_{ij} 
$$
for the $i = 1, \ldots, 9$ weeks and $j = 1, \dots, 48$ pigs and
$$
u_i \sim N(0, \sigma^2_u); v_j \sim N(0, \sigma^2_v ); \varepsilon_{ij} \sim N(0, \sigma^2_\varepsilon)
$$
all independently. That is, we assume an overall population-average growth curve $\beta_0 + \beta_1 \mathrm{week}$ and a random pig-specific shift. In other words, the effect due to week, $u_i$, is systematic to that week and  common to all pigs. The rationale behind [this model] could be that, assuming that the pigs were measured contemporaneously, we might be concerned that week-specific random factors such as weather and feeding patterns had significant systematic effects on all pigs.

Although we might prefer an alternative model,[^pig-alternative] we think that this is a reasonable specification.

[^pig-alternative]: These are repeated-measures data, which would be more conventionally modeled with autocorrelated errors within pigs. The \code{lme()} function in the \pkg{nlme} package, for example, is capable of fitting a mixed-model of this form.

The \proglang{Stata} manual fits the mixed model by maximum likelihood (rather than REML), and we duplicate the results reported there using \code{lmer()}:
```{r pigs-lmer}
m.p <- lmer(weight ~ week + (1 | id) + (1 | week),
            data=Pigs, REML=FALSE, # i.e., ML
            control=lmerControl(optimizer="bobyqa"))
summary(m.p)
```
We opt for the non-default \code{"bobyqa"} optimizer because it provides more numerically stable results for subsequent cross-validation in this example.

We can then cross-validate the model by omitting folds composed of pigs, folds composed of weeks, or folds composed of pig-weeks (which in the \code{Pigs} data set correspond to individual cases, using only the fixed effects):
```{r pigs-cv}
cv(m.p, clusterVariables="id")

cv(m.p, clusterVariables="week")

cv(m.p, clusterVariables=c("id", "week"), k=10, seed=8469)
```
We can also cross-validate the individual cases taking account of the random effects (employing the same 10 folds):
```{r pigs-cv-cases}
cv(m.p, k=10, seed=8469)
```
Because these predictions are based on BLUPs, they are more accurate than the predictions based only on fixed effects.[^crossed-vs-nested] As well, the difference between the MSE computed for the model fit to the full data and the CV estimates of the MSE is greater here than for cluster-based predictions.

[^crossed-vs-nested]: Even though there is only one observation per combination of pigs and weeks, we can use the BLUP for the omitted case because of the crossed structure of the random effects; that is each pig-week has a pig random effect and a week random effect. Although it probably isn't sensible, we can imagine a mixed model for the pig data that employs nested random effects, which would be specified by \code{lmer(weight ~ week + (1 | id/week), data=Pigs)}---that is, a random intercept that varies by combinations of \code{id} (pig) and \code{week}. This model can't be fit, however: With only one case per combination of \code{id} and \code{week}, the nested random-effect variance is indistinguishable from the case-level variance.

# Replicating cross-validation

Assuming that the number of cases $n$ is a multiple of the number of folds $k$---a slightly simplifying assumption---the number of possible partitions of cases into folds is $\frac{n!}{[(n/k)!]^k}$, a number that grows very large very quickly. For example, for $n = 10$ and $k = 5$, so that the folds are each of size $n/k = 2$, there are $113,400$ possible partitions; for $n=100$ and $k=5$, where $n/k = 20$, still a small problem, the number of possible partitions is truly astronomical, $1.09\times 10^{66}$.

Because the partition into folds that's employed is selected randomly, the resulting CV criterion estimates are subject to sampling error. (An exception is LOO cross-validation, which is not at all random.) To get a sense of the magnitude of the sampling error, we can repeat the CV procedure with different randomly selected partitions into folds. All of the CV functions in the \pkg{cv} package are capable of repeated cross-validation, with the number of repetitions controlled by the \code{reps} argument, which defaults to \code{1}.

Here, for example, is 10-fold CV for the Mroz logistic regression, repeated 5 times:

```{r mroz-reps}
cv(m.mroz, criterion=BayesRule, seed=248, reps=5, 
   method="Woodbury")
```
When \code{reps} > \code{1}, the result returned by \code{cv()} is an object of class \code{"cvList"}---literally a list of \code{"cv"} objects. The results are reported for each repetition and then averaged across repetitions, with the standard deviations of the CV criterion and the biased-adjusted CV criterion given in parentheses. In this example, there is therefore little variation across repetitions, increasing our confidence in the reliability of the results.

Notice that the seed that's set in the \code{cv()} command pertains to the first repetition and the seeds for the remaining repetitions are then selected pseudo-randomly.[^replicates] Setting the first seed, however, makes the entire process easily replicable, and the seed for each repetition is stored in the corresponding element of the \code{"cvList"} object (which isn't, however, saved in the example).

[^replicates]: Because of the manner in which the computation is performed, the order of the replicates in the \code{"cvList"} object returned by \code{cv()} isn't the same as the order in which the replicates are computed. Each element of the result, however, is a \code{"cv"} object with the correct random-number seed saved, and so this technical detail can be safely ignored. The individual \code{"cv"} objects are printed in the order in which they are stored rather than the order in which they are computed.

It's also possible to replicate CV when comparing competing models via the \code{cv()} method for \code{"modList"} objects. Recall our comparison of polynomial regressions of varying degree fit to the \code{Auto} data; we performed 10-fold CV for each of 10 models. Here, we replicate that process 5 times for each model and graph the results:
```{r model-comparison-with-reps}
#| out.width = "100%",
#| fig.height = 5,
#| fig.cap = " Replicated cross-validated 10-fold CV as a function of polynomial degree, $p$"
cv.auto.reps <- cv(models(m.1, m.2, m.3, m.4, m.5,
                        m.6, m.7, m.8, m.9, m.10),
                 data=Auto, seed=8004, reps=5)
plot(cv.auto.reps)
```
The graph shows both the average CV criterion and its range for each of the competing models.

# Cross-validating model selection

@@

Also by default, when the number of folds is 10 or fewer, \code{cvSelect()} saves the coefficients of the selected models. In this example, the \code{compareFolds()} function reveals that the variables retained by the model-selection process in the several folds are quite different:
```{r compare-selected-models}
compareFolds(cv.select)
```
## Mroz's logistic regression revisited

For a contrasting example we apply model selection to Mroz's logistic regression for married women's labor-force participation. First, recall the logistic regression model that we fit to the \code{Mroz} data:
```{r recall-Mroz-regression}
summary(m.mroz)
```

Applying stepwise model selection Mroz's logistic regression, using BIC as the model-selection criterion (via the argument \code{k=log(nrow(Mroz))} to \code{stepAIC()}) selects 5 of the 7 original predictors:
```{r mroz-selection}
m.mroz.sel <- stepAIC(m.mroz, k=log(nrow(Mroz)),
                      trace=FALSE)
summary(m.mroz.sel)
BayesRule(Mroz$lfp == "yes",
          predict(m.mroz.sel, type="response"))
```
Bayes rule applied to the selected model misclassifies 32% of the cases in the \code{Mroz} data.

Cross-validating the selected model produces a similar, slightly larger, estimate of misclassification, about 33%:
```{r cv-mroz-regression}
cv(m.mroz.sel, criterion=BayesRule, seed=345266)
```
Is this estimate of predictive performance optimistic?

We proceed to apply the model-selection procedure by cross-validation, producing more or less the same result:
```{r cv-mroz-selection}
m.mroz.sel.cv <- cvSelect(selectStepAIC, Mroz, 
                          seed=6681,
                          criterion=BayesRule,
                          model=m.mroz,
                          AIC=FALSE)
m.mroz.sel.cv
```
Setting \code{AIC=FALSE} in the call to \code{cvSelect()} uses the BIC rather than the AIC as the model-selection criterion. As it turns out, exactly the same predictors are selected when each of the 10 folds are omitted, and the several coefficient estimates are very similar, as we show using \code{compareFolds()}:
```{r compare-selected-models-mroz}
compareFolds(m.mroz.sel.cv)
```
In this example, therefore, we appear to obtain a realistic estimate of model performance directly from the selected model, because there is little added uncertainty induced by model selection.

@@

# Parallel computations

The CV functions in the \pkg{cv} package are all capable of performing parallel computations by setting the \code{ncores} argument (specifying the number of computer cores to be used) to a number > \code{1} (which is the default). Parallel computation can be advantageous for large problems, reducing the execution time of the program.

To illustrate, let's time model selection in Mroz's logistic regression, repeating the computation as performed previously and then doing it in parallel using 2 cores:
```{r parallel-computation}
system.time(m.mroz.sel.cv <- cvSelect(selectStepAIC, Mroz,
                          seed=6681,
                          criterion=BayesRule,
                          model=m.mroz,
                          AIC=FALSE))

system.time(m.mroz.sel.cv.p <- cvSelect(selectStepAIC, Mroz,
                          seed=6681,
                          criterion=BayesRule,
                          model=m.mroz,
                          AIC=FALSE,
                          ncores=2))
all.equal(m.mroz.sel.cv, m.mroz.sel.cv.p)
```

In this small problem, the parallel computation is actually *slower*, because there is an overhead cost to parallelization, but we can see that it produces the same result as before.

@@

## Adding a cost criterion

A cost criterion suitable for use with \code{cv()} or \code{cvSelect()} should take two arguments, \code{y} (the observed response vector) and \code{yhat} (a vector of fitted or predicted response values), and return a numeric index of lack of fit. The \pkg{cv} package supplies several such criteria: \code{mse(y, yhat)}, which returns the mean-squared prediction error for a numeric response; \code{rmse(y, yhat)}, which returns the (square-)root mean-squared error; \code{medAbsErr(y, yhat)}, which returns the median absolute error; and \code{BayesRule(y, yhat)} (and its non-error-checking version, \code{BayesRule2(y, yhat))}, suitable for use with a binary regression model, where \code{y` is the binary response coded \code{0} for a "failure" or \code{1} for a "success"; where \code{yhat} is the predicted probability of success; and where the proportion of *incorrectly* classified cases is returned.

To illustrate using a different prediction cost criterion, we'll base a cost criterion on the area under the receiver operating characteristic ("ROC") curve for a logistic regression. The ROC curve is a graphical representation of the classification power of a binary regression model, and the area under the ROC curve ("AUC"), which varies from 0 to 1, is a common summary measure based on the ROC [see @Wikipedia-ROC:2023]. The \pkg{Metrics} package [@HamnerFrasco:2018] includes a variety of measures useful for model selection, including an \code{auc()} function. We convert the AUC into a cost measure by taking its complement:

```{r AUCcomp}
AUCcomp <- function(y, yhat) 1 - Metrics::auc(y, yhat)
```

We then apply \code{AUCcomp()} to the the Mroz logistic regression discussed in the main \pkg{cv} package vignette, which we reproduce here, using the \code{Mroz} data frame from the \pkg{carData} package [@FoxWeisberg:2019]:
```{r Mroz-regression}
data("Mroz", package="carData")
m.mroz <- glm(lfp ~ ., data=Mroz, family=binomial)
summary(m.mroz)

AUCcomp(with(Mroz, as.numeric(lfp == "yes")), fitted(m.mroz))
```
Cross-validating this cost measure is straightforward:
```{r Mroz-CV-ROC}
library("cv")
cv(m.mroz, criterion=AUCcomp, seed=3639)
```
As expected, the cross-validated complement to the AUC is somewhat less optimistic than the criterion computed from the model fit to the whole data set.

As we explain in the vignette "Cross-validation of regression models," the \code{cv()} function differentiates between CV criteria that are averages of casewise components and criteria that are not. Computation of bias corrections and confidence intervals is limited to the former. We show in the appendix to this vignette that the AUC, and hence its complement, cannot be expressed as averages of casewise components.

\code{cv()} looks for a \code{"casewise loss"} attribute of the value returned by a CV criterion function. If this attribute exists, then the criterion is treated as the mean of casewise components, and \code{cv()} uses the unexported function \code{getLossFn()} to construct a function that returns the casewise components of the criterion.

We illustrate with the \code{mse()}:
```{r}
mse

cv:::getLossFn(mse(rnorm(100), rnorm(100)))
```
For this scheme to work, the "casewise loss" attribute must be a character string (or vector of character strings), here \code{"(y - yhat)\^2"}, that evaluates to an expression that is a function of \code{y} and \code{yhat}, and that computes the vector of casewise components of the CV criterion.

@@

### Mixed-effects models

Adding a \code{cv()} method for a mixed-model class is somewhat more complicated. We provide the \code{cvMixed()} function to facilitate this process, and to see how that works, consider the \code{"lme"} method from the \pkg{cv} package:
```{r cv.lme}
cv:::cv.lme
```
Notice that \code{cv.lme()} sets up a call to \code{cvMixed()}, which does the computational work.

Most of the arguments of \code{cvMixed()} are familiar:

* \code{model} is the mixed-model object, here of class \code{"lme"}.

* \code{package} is the name of the package in which the mixed-modeling function used to fit the model, here  \code{lme()}, resides---i.e., \code{"nlme"}; \code{cvMixed()} uses this argument to retrieve the package namespace.

* \code{data} is the data set to which the model is fit, by default extracted by the \code{get_data()} function in the \pkg{insight} package.

* \code{criterion} is the CV criterion, defaulting to the \code{mse()} function.

* \code{k} is the number of CV folds, defaulting to \code{"loo"} for CV by clusters and \code{10} for CV by cases.

* \code{reps} is the number of times the CV process is repeated, defaulting to \code{1}.

* \code{seed} is the seed for \proglang{R}'s random-number generator, defaulting to a randomly selected (and saved) value.

* \code{ncores} is the number of cores to use for parallel computation; if \code{1}, the default, then the computation isn't parallelized.

* \code{clusterVariables} is a character vector of the names of variables defining clusters; if missing, then CV is based on cases rather than clusters.

The remaining two arguments are unfamiliar:

* \code{predict.clusters.args} is a named list of arguments to be passed to the \code{predict()} function to obtain predictions for the full data set from a model fit to a subset of the data for cluster-based CV. The first two arguments should be \code{object} and \code{newdata}. It is typically necessary to tell \code{cvMixed()} how to base predictions only on fixed effects; in the case of \code{"lme"} models, this is done by setting \code{level = 0}.

* Similarly, \code{predict.cases.args} is a named list of arguments to be passed to \code{predict()} for case-based CV. Setting \code{level = 1} includes random effects in the predictions.

Finally, any additional arguments, absorbed by \code{...}, are passed to \code{update()} when the model is refit with each fold omitted. \code{cvMixed()} returns an object of class \code{"cv"}.

Now imagine that we want to support a new class of mixed-effects models. To be concrete, we illustrate with the \code{glmmPQL()} function in the \pkg{MASS} package [@VenablesRipley:2002], which fits generalized-linear mixed-effects models by penalized quasi-likelihood.[^glmmPQL] Not coincidentally, the arguments of \code{glmmPQL()} are similar to those of \code{lme()} (with an additional \code{family} argument), because the former iteratively invokes the latter; so \code{cv.glmmPQL()} should resemble \code{cv.lme()}.

[^glmmPQL]: This example is somewhat artificial in that \code{glmmPQL()} has largely been superseded by computationally superior functions, such the \code{glmer()} function in the \pkg{lme4} package. There is, however, one situation in which \code{glmmPQL()} might prove useful: to specify serial dependency in case-level errors within clusters for longitudinal data, which is not currently supported by \code{glmer()}.

As it turns out, neither the default method for \code{GetResponse()} nor \code{insight::get_data()} work for \code{"glmmPQL"} objects. These objects include a \code{"data"} element, however, and so we can simply extract this element as the default for the \code{data} argument of our \code{cv.glmmPQL()} method.

To get the response variable is more complicated: We refit the fixed part of the model as a GLM with only the regression constant on the right-hand side, and extract the response from that; because all we need is the response variable, we limit the number of GLM iterations to 1 and suppress warning messages about non-convergence:

```{r GetResponse.glmmPQL}
GetResponse.glmmPQL <- function(model, ...){
  f <- formula(model)
  f[[3]] <- 1 # regression constant only on RHS
  model <- suppressWarnings(glm(f, data=model$data, family=model$family,
                                control=list(maxit=1)))
  cv::GetResponse(model)
}
```

Writing the \code{cv()} method is then straightforward:
```{r cv.glmmPQL}
cv.glmmPQL <- function(model, data = model$data, criterion = mse,
                     k, reps = 1, seed, ncores = 1, clusterVariables, ...){
  cvMixed(
    model,
    package="MASS",
    data=data,
    criterion=criterion,
    k=k,
    reps=reps,
    seed=seed,
    ncores=ncores,
    clusterVariables=clusterVariables,
    predict.clusters.args=list(object=model,
                               newdata=data,
                               level=0,
                               type="response"),
    predict.cases.args=list(object=model,
                            newdata=data,
                            level=1,
                            type="response"),
    verbose=FALSE,
    ...)
}
```
We set the argument \code{verbose=FALSE} to suppress \code{glmmPQL()}'s iteration counter when \code{cvMixed()} calls \code{update()}.

Let's apply our newly minted method to a logistic regression with a random intercept in an example that appears in \code{?glmmPQL}:
```{r glmmPQL-example}
library("MASS")
m.pql <- glmmPQL(y ~ trt + I(week > 2), random = ~ 1 | ID,
             family = binomial, data = bacteria)
summary(m.pql)
```
We compare this result to that obtained from \code{glmer()} in the \pkg{lme4} package:
```{r compare-to-lme4}
library("lme4")
m.glmer <- glmer(y ~ trt + I(week > 2) + (1 | ID),
               family = binomial, data = bacteria)
summary(m.glmer)

  # comparison of fixed effects:
car::compareCoefs(m.pql, m.glmer) 
```
The two sets of estimates are similar, but not identical

Finally, we try out our \code{cv.glmmPQL()} method, cross-validating both by clusters and by cases,
```{r cv-example-glmmPQL, cache=TRUE}
cv(m.pql, clusterVariables="ID", criterion=BayesRule)

cv(m.pql, data=bacteria, criterion=BayesRule, seed=1490)
```
and again compare to \code{glmer()}:
```{r cv-example-glmer, cache=TRUE}
cv(m.glmer, clusterVariables="ID", criterion=BayesRule)

cv(m.glmer, data=bacteria, criterion=BayesRule, seed=1490)
```

## Adding a model-selection procedure

The \code{selectStepAIC()} function supplied by the \pkg{cv} package, which is based on the \code{stepAIC()} function from the \pkg{nnet} package [@VenablesRipley:2002] for stepwise model selection, is suitable for the \code{procedure} argument of \code{cvSelect()}. The use of \code{selectStepAIC()} is illustrated in the principal vignette for the package.

We'll employ \code{selectStepAIC()} as a "template" for writing a CV model-selection procedure. To see the code for this function, type \code{cv::selectStepAIC} at the \proglang{R} command prompt, or examine the sources for the \pkg{cv} package at https://github.com/gmonette/cv (the code for \code{selectStepAIC()} is in https://github.com/gmonette/cv/blob/main/R/cvSelect.R).

Another approach to model selection is all-subsets regression. The \code{regsubsets()} function in the \pkg{leaps} package [@LumleyMiller:2020] implements an efficient algorithm for selecting the best-fitting linear least-squares regressions for subsets of predictors of all sizes, from 1 through the maximum number of candidate predictors.[^1] To illustrate the use of \code{regsubsets()}, we employ the \code{swiss} data frame supplied by the \pkg{leaps} package:

[^1]: The \code{regsubsets()} function computes several measures of model predictive performance, including the $R^2$ and $R^2$ adjusted for degrees of freedom, the residual sums of squares, Mallows's $C_p$, and the BIC. Several of these are suitable for comparing models with differing numbers of coefficients---we use the BIC below---but all necessarily agree when comparing models with the *same* number of coefficients.

```{r swiss}
library("leaps")
head(swiss)
nrow(swiss)
```
The data set includes the following variables, for each of 47 French-speaking Swiss provinces circa 1888:

* \code{Fertility}: A standardized fertility measure.
* \code{Agriculture}: The percentage of the male population engaged in agriculture.
* \code{Examination}: The percentage of draftees into the Swiss army receiving the highest grade on an examination.
* \code{Education}: The percentage of draftees with more than a primary-school education.
* \code{Catholic}: The percentage of the population who were Catholic.
* \code{Infant.Mortality}: The infant-mortality rate, expressed as the percentage of live births surviving less than a year.

Following @LumleyMiller:2020, we treat \code{Fertility} as the response and the other variables as predictors in a linear least-squares regression:
```{r swiss-lm}
m.swiss <- lm(Fertility ~ ., data=swiss)
summary(m.swiss)

cv(m.swiss, seed=8433)
```
Thus, the RMSE for the model fit to the complete data is considerably smaller than the CV estimate of the RMSE. Can we do better by selecting a subset of the predictors, taking account of the additional uncertainty induced by model selection?

First, let's apply best-subset selection to the complete data set:
```{r subset-selection}
#| fig.cap = "Selecting the best model of each size."
swiss.sub <- regsubsets(Fertility ~ ., data=swiss)
summary(swiss.sub)

(bics <- summary(swiss.sub)$bic)
which.min(bics)

car::subsets(swiss.sub, legend="topright")
```
The graph, produced by the \code{subsets()} function in the \pkg{car} package, shows that the model with the smallest BIC is the best model with 4 predictors, including \code{Agriculture}, \code{Education}, \code{Catholic}, and \code{Infant.Mortality}, but not \code{Examination}:
```{r best-model}
m.best <- update(m.swiss, . ~ . - Examination)
summary(m.best)

cv(m.best, seed=8433) # use same folds as before
```
The RMSE for the selected model is (of course) slightly higher than for the full model fit previously, but the cross-validated RMSE is a bit lower; as we explain in the main vignette, however, it isn't kosher to select and cross-validate a model on the same data.

Here's a function named \code{selectSubsets()}, meant to be used with \code{cvSelect()}, suitable for cross-validating the model-selection process:
```{r selectSubsets}
selectSubsets <- function(data=insight::get_data(model), 
                          model,
                          indices,
                          criterion=mse,
                          save.coef=TRUE, ...){

  if (inherits(model, "lm", which=TRUE) != 1)
    stop("selectSubsets is appropriate only for 'lm' models")

  y <- GetResponse(model)
  formula <- formula(model)
  X <- model.matrix(model)

  if (missing(indices)) {
    # select the best model from the full data by BIC
    sel <- leaps::regsubsets(formula, data=data, ...)
    bics <- summary(sel)$bic
    best <- coef(sel, 1:length(bics))[[which.min(bics)]]
    x.names <- names(best)
    # fit the best model; intercept is already in X, hence - 1:
    m.best <- lm(y ~ X[, x.names] - 1) 
    fit.all <- predict(m.best, newdata=data)
    return(criterion(y, fit.all)) # return the CV criterion
  }

  # select the best model omitting the i-th fold (given by indices)
  sel.i <- leaps::regsubsets(formula, data[-indices, ], ...)
  bics.i <- summary(sel.i)$bic
  best.i <- coef(sel.i, 1:length(bics.i))[[which.min(bics.i)]]
  x.names.i <- names(best.i)
  m.best.i <- lm(y[-indices] ~ X[-indices, x.names.i] - 1)
              # predict() doesn't work here:
  fit.all.i <- as.vector(X[, x.names.i] %*% coef(m.best.i))
  fit.i <- fit.all.i[indices]
  # return the fitted values for i-th fold, CV criterion for all cases, 
  #   and the regression coefficients
  list(fit.i=fit.i, # fitted values for i-th fold
       crit.all.i=criterion(y, fit.all.i), # CV crit for all cases
       coefficients = if (save.coef){ # regression coefficients
         coefs <- coef(m.best.i)

         # fix coefficient names
         names(coefs) <- sub("X\\[-indices, x.names.i\\]", "",
                             names(coefs))

         coefs
       }  else {
         NULL
       }
  )
}
```
A slightly tricky point is that because of scoping issues, \code{predict()} doesn't work with the model fit omitting the $i$th fold, and so the fitted values for all cases are computed directly as $\widehat{\mathbf{y}}_{-i} = \mathbf{X} \mathbf{b}_{-i}$, where $\mathbf{X}$ is the model-matrix for all of the cases, and $\mathbf{b}_{-i}$ is the vector of least-squares coefficients for the selected model with the $i$th fold omitted.

Additionally, the command \code{lm(y[-indices] ~ X[-indices, x.names.i] - 1)}, which is the selected model with the $i$th fold deleted, produces awkward coefficient names like \code{"X[-indices, x.names.i]Infant.Mortality"}. Purely for aesthetic reasons, the command `sub("X\\[-indices, x.names.i\\]", "", names(coefs))` fixes these awkward names, removing the extraneous text, \code{"X[-indices, x.names.i]"}.

Applying \code{selectSubsets()} to the full data produces the full-data cross-validated RMSE (which we obtained previously):
%%```{r test-selectSubsets}
%%selectSubsets(model=m.swiss)
%%```
Similarly, applying the function to an imaginary "fold" of 5 cases returns the RMSE for the cases in the fold, based on the model selected and fit to the cases omitting the fold; the RMSE for all of the cases, based on the same model; and the coefficients of the selected model, which includes 4 or the 5 predictors (and the intercept):
%%```{r test-selectSubsets-fold}
%%selectSubsets(model=m.swiss, indices=seq(5, 45, by=10))
%%```

Then, using \code{selectSubsets()} in cross-validation, we get:
```{r cvSelect-selectSubsets}
(cv.swiss <- cvSelect(selectSubsets, model=m.swiss,
                      data=swiss, seed=8433)) # use same folds
```
Cross-validation shows that model selection exacts a penalty in RMSE. Examining the models selected for the 10 folds reveals that there is some uncertainty in identifying the predictors in the "best" model, with \code{Agriculture} sometimes appearing and sometimes not:
```{r best-models-by-folds}
compareFolds(cv.swiss)
```

@@

# Why the complement of AUC isn't a casewise CV criterion

Consider calculating AUC for folds in which a validation set contains $n_v$ observations. To calculate AUC in the validation set, we need the vector of prediction criteria, $\widehat{\mathbf{y}}_{v_{(n_v \times 1)}} = (\widehat{y}_1, ..., \widehat{y}_{n_v})^T$, and the vector of observed responses in the validation set, $\mathbf{y}_{v_{(n_v \times 1)}} = (y_1, \ldots, y_{n_v})^T$ with $y_i \in \{0,1\}, \; i = 1, \ldots, n_v$.

To construct the ROC curve, only the ordering of the values in 
$\mathbf{\widehat{y}}_v$ is relevant. Thus, assuming that there are no ties, and reordering observations if necessary, we can set 
$\mathbf{\widehat{y}}_v = (1, 2, \ldots, n_v)^T$.

If the AUC can be expressed as the casewise mean or sum of a function $\mathrm{cv}(\widehat{y}_i,y_i)$, where $\mathrm{cv}: \{1,2,...,n_v\}\times\{0,1\} \rightarrow [0,1]$, then
\begin{equation}
\label{eq:cw}
\tag{1}
\sum_{i=1}^{n_v} \mathrm{cv}(\widehat{y}_i,y_i) = \mathrm{AUC}(\mathbf{\widehat{y}}_v,\mathbf{y}_v)
\end{equation}
must hold for all $2^{n_v}$ possible values of $\mathbf{y}_v = (y_1,...,y_{n_v})^T$. If all $y\mathrm{s}$ have the same value, either 1 or 0, then the definition of AUC is ambiguous. AUC could be considered undefined, or it could be set to 0 if all $y$s are 0 and to 1 if all $y$s are 1. If AUC is considered to be undefined in these cases, we have $2^{n_v} - 2$ admissible values for $\mathbf{y}_v$.

Thus, equation (\ref{eq:cw}) produces either $2^{n_v}$ or $2^{n_v}-2$ constraints. Although there are only $2n_v$ possible values for the $\mathrm{cv(\cdot)}$ function, equation (\ref{eq:cw}) could, nevertheless, have consistent solutions.  We therefore need to determine whether there is a value of $n_v$ for which (\ref{eq:cw}) has no consistent solution for all admissible values of $\mathbf{y}_v$. In that eventuality, we will have shown that AUC cannot, in general, be expressed through a casewise sum.

>>>>

@@

@@

Also by default, when the number of folds is 10 or fewer, \code{cvSelect()} saves the coefficients of the selected models. In this example, the \code{compareFolds()} function reveals that the variables retained by the model-selection process in the several folds are quite different:
```{r compare-selected-models}
compareFolds(cv.select)
```

@@

@@

@@

@@
